Intentional release of native species undermines ecological stability

### 2022-09-21 rmd/supporting_info.Rmd
Update supporting_info.Rmd

Our goal is to compare temporal community dynamics across sites. However, the data are not comparable because of observation errors (e.g., different observers) and missing observations. To confront this challenge, we developed a Bayesian state-space model for three species groups separately: (i) whole community, the summed abundance of all species, (ii) enhanced species, the abundance of masu salmon, (iii) unenhanced species, the summed abundance of all species except masu salmon. A Bayesian state-space model is best suited for our analysis because it can account for observation errors while imputing missing values given the long-term trend at each site [@clarkPopulationTimeSeries2004; @keryBayesianPopulationAnalysis2012]. The model is composed of observation and state models, as described below.

@@

We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned weakly informative priors to parameters (**Table S6**). Three Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. Total MCMC interations ranged from `r min(df_mcmc_ssm$n_total_mcmc)` to `r max(df_mcmc_ssm$n_total_mcmc)` for the state-space models and was `r unique(df_mcmc_reg$n_total_mcmc)` for the regression models. The first `r unique(c(df_mcmc_ssm$n_burn, df_mcmc_reg$n_burn))` iterations were discarded as burn-in, and MCMC samples were saved every `r unique(df_mcmc_ssm$n_thin)` (state-space) and `r unique(df_mcmc_reg$n_thin)` iterations (regression) to reduce autocorrelation. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached \< 1.1 [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Table S7-S10**.

@@

# Table S1
source(here::here("code/table_param_set.R"))
kable(df_param, format = "markdown")

@@

kable(df_sp_list,
      format = "markdown")

@@

kable(df_prior,
      format = "markdown")

@@

kable(df_ssm,
      align = c(rep("l", 3), "c"),
      format = "markdown")

@@

# Figure S1-6
## load figures
source(here::here("code/figure_si_theory.R"))

@@

# Figure S7-9
## load figures
source(here::here("code/figure_si_obs_dyns.R"))
j <- 6

@@

source(here::here("code/figure_si_env.R"))

@@

source(here::here("code/figure_si_cor.R"))
```

>>>>

**Community dynamics comparison.** Our goal is to compare temporal community dynamics across sites. However, the data are not comparable because of observation errors (e.g., different observers) and missing observations. To confront this challenge, we developed a Bayesian state-space model for three species groups separately: (i) whole community, the summed abundance of all species, (ii) enhanced species, the abundance of masu salmon, (iii) unenhanced species, the summed abundance of all species except masu salmon. A Bayesian state-space model is best suited for our analysis because it can account for observation errors while imputing missing values given the long-term trend at each site [@clarkPopulationTimeSeries2004; @keryBayesianPopulationAnalysis2012]. The model is composed of observation and state models, as described below.

@@

**Species interaction.** XXX

**Model fitting.** We fitted the models to the data using JAGS version 4.1.0 through *runjags* package version 2.2.0-2 in R [@denwoodRunjagsPackageProviding2016]. We assigned weakly informative priors to parameters (**Table S6**). Three Markov chain Monte Carlo (MCMC) chains were run until parameter estimates converged. Total MCMC interations ranged from `r min(df_mcmc_ssm$n_total_mcmc)` to `r max(df_mcmc_ssm$n_total_mcmc)` for the state-space models and was `r unique(df_mcmc_reg$n_total_mcmc)` for the regression models. The first `r unique(c(df_mcmc_ssm$n_burn, df_mcmc_reg$n_burn))` iterations were discarded as burn-in, and MCMC samples were saved every `r unique(df_mcmc_ssm$n_thin)` (state-space) and `r unique(df_mcmc_reg$n_thin)` iterations (regression) to reduce autocorrelation. Convergence was assessed by examining whether the $\hat{R}$ indicator of each parameter approached \< 1.1 [@gelmanDataAnalysisUsing2007]. Data manipulation and analysis were performed in R version 4.1.0 [@rcoreteamLanguageEnvironmentStatistical2021]. Parameter estimates were summarized in **Table S7-S10**.

@@

# Table S1
source(here::here("code/table_param_set.R"))
knitr::kable(df_param, format = "markdown")

@@

knitr::kable(df_sp_list,
             format = "markdown")
```

@@

knitr::kable(df_prior,
             format = "markdown")

@@

knitr::kable(df_ssm,
             align = c(rep("l", 3), "c"),
             format = "markdown")

@@

# Figure S1-6
## load figures
source(here::here("code/si_figure_theory.R"))

@@

# Figure S7-9
## load figures
source(here::here("code/si_figure_obs_dyns.R"))
j <- 6

@@

source(here::here("code/si_figure_env.R"))

@@

source(here::here("code/si_figure_cor.R"))
```
